---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
devtools::load_all()
```

# futebolplotR 

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/futebolplotR)](https://CRAN.R-project.org/package=futebolplotR)

<!-- badges: end -->

The code for this package was copied **heavily** from [nbaplotR](https://mrcaseb.github.io/nbaplotR/) with minor changes to support Brazilian soccer team logos.

The goal of futebolplotR is to provide functions and geoms that help visualization of Brazilian soccer related analysis. It provides a ggplot2 geom that does the heavy lifting of plotting brazilian soccer logos in high quality, with correct aspect ratio and possible transparency.

## Installation

<!-- The easiest way to get futeboplotR is to install it from [CRAN](https://cran.r-project.org/package=nbaplotR) with: -->

<!-- ``` r -->
<!-- install.packages("nbaplotR") -->
<!-- ``` -->

To get a bug fix or to use a feature from the development version, you can install the development version of futebolplotR either from [GitHub](https://github.com/brunomioto/futebolplotR/) with:

``` r
if (!require("pak")) install.packages("pak")
pak::pak("brunomioto/futebolplotR")
```

## Using futebolplotR

The key function in the package is geom_futebol_logo() which will add Brazilian soccer team logos to a ggplot. Here you can see all the current available teams to plot.

```{r, plot, dpi=300, message=FALSE, warning=FALSE}
library(futebolplotR)
library(ggplot2)

team <- futebolplotR::valid_team_names()
team <- team[1:20]
df <- data.frame(
  a = rep(1:5, 4),
  b = sort(rep(1:4, 5), decreasing = TRUE),
  teams = team
)

ggplot(df, aes(x = a, y = b)) +
  geom_futebol_logos(aes(team_abbr = teams), width = 0.1) +
  geom_label(aes(label = teams), nudge_y = -0.35)+
  theme_void()+
  theme(
    plot.margin = margin(15,15,15,15,"pt")
  )+
  coord_cartesian(clip = "off")
```

