---
title: "Get Started with futebolplotR"
description:
author: "Bruno Mioto & Gabriela Junqueira"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  dpi = 600
)
```

<!-- # fazer uma intro -->

# Install the Package

The easiest way to get `FutebolplotR` is to install it from with:

```{r eval = FALSE}
if (!require("pak")) install.packages("pak")
pak::pak("brunomioto/futebolplotR")
```

# Common usage


Let's explore a few standard scenarios for utilizing `FutebolplotR`.
Initially, we load the essencial packages.


```{r setup, message=FALSE, warning=FALSE}
library(futebolplotR)
library(ggplot2)
library(dplyr)
library(brasileirao)
```

We will load the data for the Campeonato Brasileiro and explore the database.
We will select the 2023 season of the Campeonato Brasileiro


```{r}
season_2023 <- brasileirao::matches %>%
  dplyr::filter(season == 2023) %>%
  brasileirao::add_variables()

season_2023_home <- season_2023 %>%
    group_by(home) %>%
    mutate(total_home_points = sum(home_points),
           goal_concede_home = sum(as.numeric(away_score)),
           goal_scored_home = sum(as.numeric(home_score))) %>%
    ungroup() %>%
    slice(1:20) %>%
    select(home, home_abbr, total_home_points, goal_concede_home, goal_scored_home) %>%
    rename(fc = home)

season_2023_away <- season_2023 %>%
    group_by(away) %>%
    mutate(total_away_points = sum(away_points),
           goal_concede_away = sum(as.numeric(home_score)),
           goal_scored_away = sum(as.numeric(away_score))) %>%
    ungroup() %>%
    slice(1:20) %>%
    select(away, total_away_points, goal_concede_away, goal_scored_away) %>%
    rename(fc = away)


df_teams <- season_2023_home %>%
    left_join(season_2023_away, by = "fc") %>%
    group_by(fc) %>%
    mutate(goals_concede = sum(goal_concede_home, goal_concede_away),
           goals_scored = sum(goal_scored_home, goal_scored_away),
           goals_difference = goals_scored - goals_concede) %>%
    ungroup()
```

Firstly, we will visualize the points earned at home and away for each team.
This will be done to demonstrate how we can use the team badges.

```{r}
ggplot(df_teams, aes(x = total_home_points, y = total_away_points)) +
  geom_abline(color = "grey30")+
  futebolplotR::geom_futebol_badges(aes(team_abbr = home_abbr),
                                    width = 0.064, alpha = 0.8) +
  labs(
    x = "Points home",
    y = "Points away",
    caption = "Data: @brasileirao",
    title = "Campeonato Brasileiro 2023"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    plot.title.position = "plot"
  ) +
  coord_fixed(
    xlim = c(5,45),
    ylim = c(5,45)
  )
```
Now, we will demonstrate how we can incorporate each team's specific colors into the shapes (in this case, the bars of the graph)
and how we can use the team badges as axis labels.

```{r}
ggplot(df_teams, aes(x = reorder(home_abbr, -goals_scored),
                              y = goals_scored)) +
  geom_col(aes(color = home_abbr, fill = home_abbr),
           width = 0.5) +
  futebolplotR::scale_color_futebol(type = "secondary") +
  futebolplotR::scale_fill_futebol(type = "primary") +
  labs(
    title = "Campeonato Brasileiro 2023",
    y = "Goals Scored",
    caption = "Data: @brasileirao"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    plot.title.position = "plot",
    axis.title.x = element_blank(),
    #this line triggers the replacement of team abbreviations with logos
    axis.text.x = futebolplotR::element_futebol_badge(size = 0.8)
  )
```

```{r}
ggplot(df_teams, aes(x = reorder(home_abbr, -goals_difference),
                              y = goals_difference)) +
  geom_col(aes(color = home_abbr, fill = home_abbr),
           width = 0.5) +
  futebolplotR::scale_color_futebol(type = "secondary") +
  futebolplotR::scale_fill_futebol(type = "primary") +
  labs(
    title = "Compeonato Brasileiro 2023",
    y = "Goals difference"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    plot.title.position = "plot",
    axis.title.x = element_blank(),
    #this line triggers the replacement of team abbreviations with logos
    axis.text.x = futebolplotR::element_futebol_badge(size = 0.8)
  )
```
